---
title: "Reproducible Research: Peer Assessment 1"
author: Cedric Bonjour
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
library(data.table)

raw <- fread("activity.csv")

head(raw[!is.na(steps)])


```



## What is mean total number of steps taken per day?

Part 1

```{r}
t.steps <- aggregate( steps~date, raw, sum)
names(t.steps)[2] <- "total.steps"
t.steps
```

Part 2

```{r}
hist(t.steps$total.steps, 
     breaks=15,
     xlab="Steps Taken Per Day", 
     main="Histogram of Steps Taken Per Day")

```

Part 3 

Total steps per day mean and median

```{r}
c("mean" = mean(t.steps$total.steps), "median" =median(t.steps$total.steps))
```



## What is the average daily activity pattern?

Part 1

```{r}

m.interval <-aggregate( steps ~ interval, raw ,mean,na.rm=TRUE)

names(m.interval)[2] <- "meanSteps"

plot(m.interval$interval,m.interval$meanSteps,
     type="l",
     xlab="5-Minute Interval",
     ylab="Mean Steps Taken",
     main="Steps during the Day")


```
Part 2
```{r}
 max <- m.interval[  which.max(m.interval$meanSteps), ]
names(max)[2] <- "max.steps"
max
 # c("" = m.interval$interval[which.max(m.interval$steps)])
```



## Imputing missing values

Part 1
```{r}
c ("Number of rows with missing data" = sum(!complete.cases(raw)))

```

Part 2 & 3

```{r}
fill <- merge(raw , m.interval, by="interval", sort = F)
fill[is.na(steps)]$steps <- round (fill[is.na(steps), meanSteps])
fill[,meanSteps:=NULL]
fill
```
Part 4 

```{r}
t.steps.filled <- aggregate( steps~date, fill, sum)
names(t.steps.filled)[2] <- "total.steps"
t.steps.filled
```


```{r}
hist(t.steps.filled$total.steps, 
     breaks=15,
     xlab="Steps Taken Per Day", 
     main="Histogram of Steps Taken Per Day")

```
Total steps per day mean and median

```{r}
c("mean" = mean(t.steps.filled$total.steps), "median" =median(t.steps.filled$total.steps))
```


We notice that imputing on missing values does not affect the results much


## Are there differences in activity patterns between weekdays and weekends?

```{r}

fill$weekday<-weekdays(as.Date(fill$date))

fill$dayType <- factor(ifelse(fill$weekday=="Saturday"|fill$weekday=="Sunday", "weekend", "weekday"))

fill

```
```{r}
r <-aggregate(steps~interval+dayType, fill, mean)
names(r)[3] <- "mean.steps"
r
```
```{r}
library(lattice)
xyplot(r$mean.steps ~ r$interval|r$dayType,
       r,
       type="l",
       layout=c(1,2),
       xlab="Interval",
       ylab="Number of Steps")
```
